---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF - SS 2022/23"
author: "Levin Moggert (12026061)"
date: "`r Sys.Date()`"
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers
    (Martrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your
    results.
-   Use a level 2 heading for each new exercise and answer each subtask
    next to it's bullet point or use a new level 3 heading if you want.
-   Always render the R code for your solutions and make sure to include
    the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).
-   Always submit both the rendered document(s) as well as your source
    Rmarkdown document. Submit the files separately on moodle, **not**
    as a zip archive.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `exeRcise-sheet-3`. If you forgot how to do this, you can follow
    this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)
    
With git it is easier to work on projects in a group, but also just sharing the projects online in an easily accessable way is possible. On the downside the git integration in R does not automatically update changes made in the project, and if multiple people are working on the same project at the same time it can't adapt to multiple people simultaniously making changes to a document and committing them to git.
    
e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
    
    https://github.com/LevinM2101/exeRcise-sheet-3
    
b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Exercise 3: Baby-Names in Munich (4.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2022` and
`2021`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.

b)  Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.
    
```{r}
BabyN21 <- read.csv("vornamen_2021.csv")
BabyN22 <- read.csv("open_data_portal_2022.csv")

typeof(BabyN21$Anzahl)
typeof(BabyN22$Anzahl)

BabyN21$Anzahl <- as.numeric(BabyN21$Anzahl)
BabyN22$Anzahl <- as.numeric(BabyN22$Anzahl)

BabyN21$Anzahl <- ifelse(is.na(BabyN21$Anzahl), 0, BabyN21$Anzahl)
BabyN22$Anzahl <- ifelse(is.na(BabyN22$Anzahl), 0, BabyN22$Anzahl)
```
    
In both datasets if a name has been picked 4 times or less the value is "4 oder weniger", that's probably the reason it was a chr variable. I just used as.numeric because I can't know the actual value of each "4 oder weniger" and I can't use a value like "<=4" without changing the variables into chr again so I might as well just turn them into 0 for each.

c)  Calculate the total number of babies born in Munich in 2022
    and 2021. Which year had the bigger baby-boom?
    
```{r}
sum(BabyN21$Anzahl)
sum(BabyN22$Anzahl)
sum(BabyN21$Anzahl) > sum(BabyN22$Anzahl)
```
2021 probably had the bigger baby boom. 

d)  Add a new column `year` to both datasets which holds the correct
    year for each.
    
```{r}
library(dplyr)

BabyN21 <- BabyN21 %>% mutate(year = 2021)
BabyN22 <- BabyN22 %>% mutate(year = 2022)
```
    

e)  Combine both datasets into one using `bind_rows()`.

```{r}
BabyNames <- bind_rows(BabyN21, BabyN22)
```


f)  Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.
    
```{r}
library(knitr)

TopBabyNames <- BabyNames %>% group_by(BabyNames$ï..Vorname) %>% summarize(total = sum(Anzahl)) %>% arrange(desc(total))
Top10Names <- TopBabyNames %>% top_n(10, total)
kable(Top10Names, caption = "Top 10 names munich 2021-2022")
```
    

## Exercise 4: Chat GPT + apply (3 points)

For this task: Specifically use ChatGPT to solve the task and submit
your prompts in addition to the solution

a)  The code below does not work because the wrong apply function has
    been used. Find out which apply function would be correct and why it
    did not work. Correct the code. Also calculate the rowwise means.

###Create a sample data frame

tax_data \<- data.frame( Name = c("Munich GmbH", "ABC Inc.", "Backpacks
1980", "Bavarian Circus"), Tax_2019 = c(5000, 4000, 6000, 3500),
Tax_2020 = c(4800, 4200, 5800, 3700), Tax_2021 = c(5200, 3800, 5900,
3400) )

###Calculate column-wise means

column_means \<- lapply(tax_data\[, -1\], 2, mean)

column_means

This code doesn't work because the wrong apply function has been used: tax_data \<- data.frame( Name = c("Munich GmbH", "ABC Inc.", "Backpacks
1980", "Bavarian Circus"), Tax_2019 = c(5000, 4000, 6000, 3500),
Tax_2020 = c(4800, 4200, 5800, 3700), Tax_2021 = c(5200, 3800, 5900,
3400) ) column_means \<- lapply(tax_data\[, -1\], 2, mean)
 Find out which apply function would be correct and why it did not work. Correct the code. Also calculate the rowwise means.
 
```{r}
# Create the tax_data data frame
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)

# Calculate the column means using sapply
column_means <- sapply(tax_data[, -1], mean)

# Add the row-wise means as a new column
tax_data$Row_Means <- rowMeans(tax_data[, -1])

# Print the updated tax_data data frame and the column means
tax_data
column_means
```


b)  Using ChatGPT try to understand what the rapply() function does.
    Create an easy example with mock data where the function is used and
    explain it in your words.
    
In r create an easy example with mock data where the rapply() function is used and explain to me what the function does.

```{r}
# Create a nested list
nested_list <- list(a = list(1, 2), b = list(3, 4), c = list(5, list(6, 7)))

# Print the original list
print(nested_list)

# Use rapply to increment each element by 1
result <- rapply(nested_list, function(x) x + 1)

# Print the result
print(result)
```

The rapply() function performs an operation on each element of a list, even if they're in a list within a list. In this example there is 1 added to every element.

## Final Note

Make sure to push all your commits and changes to GitHub before
submittining the exercise sheet.
